---
title: "R Code"
author: '540741586'
date: "2025-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load required packages
# install.packages("tidyverse")
# install.packages("car")
# install.packages("lmtest")
# install.packages("effects")
library(tidyverse)
library(car)
library(lmtest)
library(effects)

# Load dataset from GitHub
temp <- tempfile()
url <- "https://github.com/BORUI99/DATA2002-Project/raw/main/wine%2Bquality.zip"
download.file(url, temp, mode = "wb")
unzip(temp, exdir = tempdir())
wine_red <- read.csv(file.path(tempdir(), "winequality-red.csv"), sep = ";")

summary(wine_red)
colSums(is.na(wine_red))   # no missing values

# Fit full multiple linear regression (MLR) model
full.model <- lm(quality ~ ., data = wine_red)
summary(full.model)

# Linearity & Homoscedasticity
par(mfrow = c(1, 2))
plot(full.model$fitted.values, resid(full.model),
     main = "Residuals vs Fitted", xlab = "Fitted", ylab = "Residuals")
abline(h = 0, col = "red")

# Breusch–Pagan test to check for heteroscedasticity
bp <- bptest(full.model)
bp$p.value   # p-value ≈ 1.6e-09 meaning some heteroscedasticity

# Normality
qqnorm(rstandard(full.model), main = "Q–Q Plot of Standardized Residuals")
qqline(rstandard(full.model), col = "red")

# Collinearity
vif(full.model)   # all < 2 so they are stable coefficients

# Backward AIC variable selection
step.back.aic <- step(full.model, direction = "backward")

# Compare models
AIC(full.model, step.back.aic)  # ≈ 3162.28 vs 3156.98
summary(step.back.aic)$r.squared       # ≈ 0.3595
summary(step.back.aic)$adj.r.squared   # ≈ 0.3567

# Check significance of dropped terms
drop1(step.back.aic, test = "F")

# Predicted vs Observed plot
pred <- predict(step.back.aic)
plot(pred, wine_red$quality,
     main = "Predicted vs Observed Quality",
     xlab = "Predicted", ylab = "Observed")
abline(0, 1, col = "red", lwd = 2)


# Standardized coefficients for relative effect size
wine_scaled <- as.data.frame(scale(wine_red))
std.model <- lm(quality ~ ., data = wine_scaled)
summary(std.model)

# Extract significant variables
coef(summary(step.back.aic))

# Partial dependence plots for interpretability
plot(allEffects(step.back.aic))

# Key findings:
# - Alcohol: β ≈ +0.289 → higher alcohol = higher quality
# - Volatile acidity: β ≈ −1.013 → higher VA = lower quality
# - Sulphates, pH: smaller positive/negative effects
# - R² ≈ 0.36, Adj-R² ≈ 0.357 → good fit with fewer terms
```
#D
library(tidyverse)
library(corrplot)
library(ggplot2)
library(gridExtra)
library(readr)
library(tidyverse)
library(caret)
library(randomForest)
library(xgboost)
library(pROC)
library(PRROC)
library(pdp)
library(ggplot2)
library(gridExtra)
library(patchwork)
library(yardstick)
#PR curve and threshold selection
pr_curve = pr.curve(scores.class0 = predictions[y_test == 1],
                     scores.class1 = predictions[y_test == 0],
                     curve = TRUE)
pr_data = data.frame(
  Recall = pr_curve$curve[, 1],
  Precision = pr_curve$curve[, 2],
  Threshold = pr_curve$curve[, 3]
)

# Find the optimal F1 threshold
f1_scores <- 2 * (pr_data$Precision * pr_data$Recall) / (pr_data$Precision + pr_data$Recall)
best_f1_idx <- which.max(f1_scores)
best_threshold <- pr_data$Threshold[best_f1_idx]

pr_plot <- ggplot(pr_data, aes(x = Recall, y = Precision)) +
  geom_line(color = "darkgreen", linewidth = 1.2) +
  geom_vline(xintercept = pr_data$Recall[best_f1_idx], 
             linetype = "dashed", color = "red") +
  geom_hline(yintercept = pr_data$Precision[best_f1_idx], 
             linetype = "dashed", color = "red") +
  geom_point(data = pr_data[best_f1_idx, ], 
             aes(x = Recall, y = Precision), 
             color = "red", size = 3) +
  geom_text(aes(x = pr_data$Recall[best_f1_idx] + 0.1, 
                y = pr_data$Precision[best_f1_idx] - 0.1),
            label = paste("threshold value =", round(best_threshold, 3)),
            color = "red", size = 4) +
  labs(title = "Precision - recall curve (PR Curve)",
       x = "Recall",
       y = "Precision") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```{r}
# S15: Error analysis and sensitivity
cat("\n=== S15: Error analysis and sensitivity ===\n")

# Re-predict using the optimal threshold
final_predictions <- ifelse(predictions > best_threshold, 1, 0)

# confusion matrix
conf_matrix <- table(
  Predicted = factor(final_predictions, levels = c(0, 1)),
  Actual = factor(y_test, levels = c(0, 1))
)

cat("confusion matrix:\n")
print(conf_matrix)

# Calculate the indicators of various categories
conf_matrix_df <- as.data.frame(conf_matrix)
colnames(conf_matrix_df) <- c("Predicted", "Actual", "Count")

conf_matrix_plot <- ggplot(conf_matrix_df, aes(x = Actual, y = Predicted, fill = Count)) +
  geom_tile(alpha = 0.8) +
  geom_text(aes(label = Count), color = "white", size = 6) +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "confusion matrix",
       x = "True category",
       y = "Class prediction") +
  theme_minimal()

conf_matrix_plot

# Analysis of Error Cases
test_data_with_pred <- wine_data_clean[-train_index, ] %>%
  mutate(
    predicted = final_predictions,
    probability = predictions,
    error_type = case_when(
      predicted == 1 & quality_binary == 0 ~ "False Positive",
      predicted == 0 & quality_binary == 1 ~ "False Negative",
      predicted == quality_binary ~ "Correct"
    )
  )

# Analysis of Error Cases
error_analysis <- test_data_with_pred %>%
  filter(error_type != "Correct") %>%
  group_by(error_type) %>%
  summarise(
    count = n(),
    avg_alcohol = mean(alcohol),
    avg_volatile_acidity = mean(volatile.acidity),
    avg_sulphates = mean(sulphates),
    avg_citric_acid = mean(citric.acid)
  )

cat("\nAnalysis of Error Cases:\n")
print(error_analysis)

# Save the first 20 error cases
error_cases_top20 <- test_data_with_pred %>%
  filter(error_type != "Correct") %>%
  arrange(desc(abs(probability - 0.5))) %>%
  head(20) %>%
  select(fixed.acidity:alcohol, quality, quality_binary, predicted, probability, error_type)



# Threshold sensitivity analysis
thresholds <- seq(0.3, 0.7, 0.05)
threshold_analysis <- map_df(thresholds, function(thresh) {
  pred_class <- ifelse(predictions > thresh, 1, 0)
  conf_matrix <- table(factor(pred_class, levels = c(0,1)), 
                       factor(y_test, levels = c(0,1)))
  
  if(nrow(conf_matrix) == 2 && ncol(conf_matrix) == 2) {
    precision <- conf_matrix[2,2] / sum(conf_matrix[2,])
    recall <- conf_matrix[2,2] / sum(conf_matrix[,2])
    f1 <- 2 * (precision * recall) / (precision + recall)
    accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
  } else {
    precision <- recall <- f1 <- accuracy <- 0
  }
  
  data.frame(
    Threshold = thresh,
    Precision = precision,
    Recall = recall,
    F1 = f1,
    Accuracy = accuracy
  )
})

# Threshold sensitivity map
threshold_plot <- threshold_analysis %>%
  pivot_longer(cols = c(Precision, Recall, F1, Accuracy), 
               names_to = "Metric", values_to = "Value") %>%
  ggplot(aes(x = Threshold, y = Value, color = Metric)) +
  geom_line(linewidth = 1.2) +
  geom_vline(xintercept = best_threshold, linetype = "dashed", color = "red") +
  geom_point(data = threshold_analysis[threshold_analysis$Threshold == best_threshold, ] %>% 
               pivot_longer(cols = c(Precision, Recall, F1, Accuracy), 
                            names_to = "Metric", values_to = "Value"),
             aes(x = Threshold, y = Value), size = 3) +
  labs(title = "Threshold sensitivity analysis",
       x = "classification threshold",
       y = "indexscore") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

threshold_plot
```
